### 2017-06-19 sections/implementation.tex
Update on Overleaf.

SVD decomposition is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results form the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements. As an example can serve temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc. There are two similar sets of results. One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored. In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled. The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored. Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields. It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.

@@

Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$.

>>>>

SVD decomposition is applied on matrices. The first thing to do is therefore to assemble an input matrix. Results from the finite element method are scalar, vector or tensor fields represented by discrete values calculated in nodes of the mesh or in integration points on finite elements. As an example can serve temperature field, vector of nodal displacements, strain tensor evaluated in integration points, etc. There are two similar sets of results. One is generated by a non-linear algorithms, where several incremental steps are stored and the other is generated by time integration, where results in particular time steps are stored. In order to compress data, an auxiliary matrix $\mtrx{A}$ has to be assembled. The number of rows of the matrix $\mtrx{A}$ is equal to the number of incremental or time steps while the number of columns is equal to the number of points in which the results are stored. Such auxiliary matrix is assembled for each scalar field and for each component of the vector and tensor fields. It means, three matrices corresponding to the displacement in the $x$, $y$ and $z$ directions are assembled for the vector of displacements in three-dimensional problems.

@@

Using formulas \eqref{eq:svd-expansion} and \eqref{eq:svd-approx-expansion} the equation \eqref{eq:elem-sqr-sigma-sqr} can be applied to the difference between original matrix $\mtrx{A}$ and approximation matrix $\mtrx{A'}$

### 2017-06-19 sections/math.tex
Update on Overleaf.

\begin{equation}
\mtrx{A}=\sum_{i=1}^{n} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}},
\label{eq:svd-expansion}
\end{equation}

\noindent
where $s_i$ is the $i$-th singular value, $\bf{u}_i$ and $\bf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$. Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ... \geq s_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$. Therefore, if we take only first $r$ members of the above summation we get an approximation matrix

\begin{equation}
\mtrx{A'}=\sum_{i=1}^{r} s_{i}\bf{u}_i\bf{v}_i^{\mathsf{T}}.
\label{eq:svd-approx-expansion}
\end{equation}

@@

\subsection{Randomized SVD}

The exact SVD of a $m \times n$ matrix has computational complexity \newline $O(min(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

As the implementation of the randomized SVD algorithm RedSVD library was used. RedSVD is an open-source software package for solving several matrix decompositions. The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}. The brief summary of the algorithm steps follows below.

% redsvd algorithm description
Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then the result is full SVD). As the first step, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed. Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized. Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed. Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$, i.e. $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$. Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.

Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed. As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$. Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.

Finally, the algorithm performs SVD of $\mtrx{C}$ using traditional exact SVD solver and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$, where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$. Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.

$\mtrx{A}$ is then decomposed as

\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
\\
&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
\\
&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
\label{eq:redsvd}
\end{eqnarray}

\noindent
Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$. $\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.

% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$. Note that $\mtrx{AA^T}$ is the dataset covariance matrix. Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$. Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.

>>>>

\begin{equation}
\mtrx{A}=\sum_{i=1}^{n} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}},
\label{eq:svd-expansion}
\end{equation}

\noindent
where $s_i$ is the $i$-th singular value, $\mathbf{u}_i$ and $\mathbf{v}_i$ are corresponding singular vectors, and $n$ is the rank of matrix~$\mtrx{A}$. Considering the fact that singular values are ordered $s_{1} \geq s_{2} \geq s_{3} \geq ... \geq s_{n}$, the above formula implies that the first term of the sum would have the highest contribution and the last term would have the lowest contribution to matrix~$\mtrx{A}$. Therefore, if we take only first $r$ members of the above summation we get an approximation matrix

\begin{equation}
\mtrx{A'}=\sum_{i=1}^{r} s_{i}\mathbf{u}_{i}\mathbf{v}_{i}^{\mathsf{T}}.
\label{eq:svd-approx-expansion}
\end{equation}

@@

% subsection Randomized SVD
\input{sections/randomized-svd.tex}
